---
title: "Data Exploration"
output: 
  html_notebook: 
    number_sections: yes
    toc: yes
---

# Introduction
In this notebook I carry out basic exploration of the data set. Observations and notes made here will be used to inform feature generation, source code creation, and model selection. 

The ultimate goal of this project is to develop a machine learning model that can accurately identify online gamblers prone to addiction based on their gambling behavior. The data source used for this project is [Behavioral Characteristics of Internet Gamblers Who Trigger Corporate Responsible Gambling Interventions](http://www.thetransparencyproject.org/datasets) and is included in the repository. The details of the data set are documented in the [Codebook](../references/Code Book.pdf)

# Data Load & Inspection
```{r setup, collapse=TRUE}
library(readr)
library(tibble)
library(dplyr)
library(magrittr)
library(ggplot2)
```

```{r, collapse=TRUE}
df <- read_csv(file.path("..", 'data', 'raw', 'gambling.csv'))

print(dim(df))
df %>% 
    as_tibble() %>% 
    head()
```

